
---
title: "EMBL Career Path Study"
author: "Junyan Lu and Rachel Coulthard-Graf"
date: "Last updated on `r doc_date()`"
output: 
    BiocStyle::html_document:
      toc_float: true
---

<font size="4">This document contains the codes that generate the data analysis results and figures presented in the manuscript "PhD and postdoc training outcomes at EMBL: changing career paths for life scientists in Europe".</font> 

# Section 1: preparation 

## Load required R packages
```{r, warning=FALSE, message=FALSE}
library(gridExtra)
library(cowplot)
library(ggbeeswarm)
library(Hmisc)
library(ggpubr)
library(reshape2)
library(psych)
library(plotly)
library(orca)
library(RColorBrewer)
library(survival)
library(glmnet)
library(survminer)
library(tidyverse)
library(reticulate)
#load helper functions
source("code/utils.R")

#silent warnings and messages
knitr::opts_chunk$set(warning = FALSE, message = FALSE, autodep = TRUE)
```

## Read and pre-process dataset

Due to date protection, the original dataset will not be included in the public version. 
```{r}
#original data table
originalData = read_csv("data/datafinal_forR_050222.csv")
```

### Pre-processing

Years as integer
```{r}
careerData <- mutate_at(originalData, 
                        vars(`from.year`, to_year,phd_year_if_known),
                        as.integer) %>% rename(from_year = `from.year`)
```

Calculate Year at EMBL
```{r}
careerData <- mutate(careerData, year_at_EMBL = to_year - from_year)
```

Change unknown to NA in all character columns
```{r}
subStr <- function(x) {
  x[x %in% c("unknown","Unknown")] <- NA
  x
}
careerData <- mutate_if(careerData, is.character, subStr)
```

#### Some additonal changes of the variable names

Change CV_wasNonR to CV_wasNonRes
```{r}
careerData <- dplyr::rename(careerData, CV_wasNonRes = CV_wasNonR)
```

Change PhDtoGL to PhDtoPI
```{r}
careerData <- dplyr::rename(careerData, PhDtoPI = PhDtoGL)
```


# Section 2: Overview of the career outcome

## Career outcomes for EMBL PhD and postdoc alumni

### Prepare data for plotting
```{r}
tempdata <- originalData
tempdata$timepointLast <- as.character(tempdata$timepointLast)
tempdata$timepointLast <- gsub('AcFac', 'AcPI', tempdata$timepointLast)
tempdata$timepointLast <- factor(tempdata$timepointLast, 
                                 levels = c("unknown", "NonSci", "NonRes", "IndR", "AcPD", "AcOt", "AcPI"))
tempdata$gender <- as.factor(tempdata$gender)
predoc <- tempdata[ which(tempdata$type_pre_postdoc=='predoc'),]
postdoc <- tempdata[ which(tempdata$type_pre_postdoc=='postdoc'),]
tempdata$type_pre_postdoc <- factor(tempdata$type_pre_postdoc, levels = c("predoc", "postdoc"))
##change colours###
mycols<-setNames(c("#D0D0CE","#E2E868","#F4C61F","#F49E17","#D41645","#734595","#3B6FB6"),
                 c('unknown','NonSci','NonRes','IndR','AcPD','AcOt', 'AcPI'))
```

### Plot (Figure 1B)

```{r figure1B, fig.width=8, fig.height=4, dpi=200}
#predoc
predoc_cohort <- predoc %>%
  dplyr::group_by(gender, timepointLast) %>%
  dplyr::summarize(count = n())

a <- ggplot(predoc_cohort, aes(fill=timepointLast, y=count, x=gender),col=mycols) + 
  geom_bar( stat="identity", position="fill") + scale_fill_manual(values=c(mycols)) + 
  labs(x = "Gender", y="Proportion in each role") + 
  theme_minimal(base_size = 16)+ ggtitle("PhD students,
current role") + 
  theme(axis.title=element_text(size=12))+ theme(plot.title=element_text(size=12)) 


#postdoc
postdoc_cohort <- postdoc %>%
  dplyr::group_by(gender, timepointLast) %>%
  dplyr::summarise(count = n())

ep <- ggplot(postdoc_cohort,  aes(fill=timepointLast, y=count, x=gender),col=mycols) + 
  geom_bar( stat="identity", position="fill") + 
  scale_fill_manual(values=c(mycols)) + 
  labs(x = "Gender", y="Proportion in each role") + 
  theme_minimal(base_size = 16) + ggtitle("Postdocs, 
current role") + 
  theme(axis.title=element_text(size=12)) + theme(plot.title=element_text(size=12))+
  theme(legend.key.height = unit(1, 'lines'),legend.key.width = unit(0.4, 'lines'),legend.title = element_text(size=12), legend.text = element_text(size=12)) 


# make legend
g_legend <- function(a.gplot){
  
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
  
  }

mylegend <- g_legend(ep)

#combine plots
grid.arrange(arrangeGrob(a + theme(legend.position="none"),
                               ep + theme(legend.position="none"),
                               mylegend,
                               nrow=1,widths=c(4,4,3)))
```
    
## Positions stratified by years after EMBL 

### Prepare data for plotting
```{r}
mydata2 <- tempdata[ which(tempdata$completeness=='Complete CV'),]

mydata2$cohort <- factor(mydata2$cohort, levels = c("1997-2004","2005-2012","2013-2020"))
mydata2$position_1 <- as.character(mydata2$position_1)
mydata2$position_5 <- as.character(mydata2$position_5)
mydata2$position_9 <- as.character(mydata2$position_9)
mydata2$position_13 <- as.character(mydata2$position_13)
mydata2$position_17 <- as.character(mydata2$position_17)
mydata2$position_1 <- gsub('laterPI', 'CV-gap', mydata2$position_1)
mydata2$position_5 <- gsub('laterPI', 'CV-gap', mydata2$position_5)
mydata2$position_9 <- gsub('laterPI', 'CV-gap', mydata2$position_9)
mydata2$position_13 <- gsub('laterPI', 'CV-gap', mydata2$position_13)
mydata2$position_17 <- gsub('laterPI', 'CV-gap', mydata2$position_17)
mydata2$timepointLast <- gsub('laterPI', 'CV-gap', mydata2$timepointLast)
mydata2$position_1 <- gsub('AcFac', 'AcPI', mydata2$position_1)
mydata2$position_5 <- gsub('AcFac', 'AcPI', mydata2$position_5)
mydata2$position_9 <- gsub('AcFac', 'AcPI', mydata2$position_9)
mydata2$position_13 <- gsub('AcFac', 'AcPI', mydata2$position_13)
mydata2$position_17 <- gsub('AcFac', 'AcPI', mydata2$position_17)
mydata2$timepointLast <- gsub('AcFac', 'AcPI', mydata2$timepointLast)
mydata2$position_1 <- factor(mydata2$position_1, levels = c("","CV-gap", "NonSci", "NonRes", "IndR", "AcPD", "AcOt", "AcPI"))
mydata2$position_5 <- factor(mydata2$position_5, levels = c("","CV-gap", "NonSci", "NonRes", "IndR", "AcPD", "AcOt", "AcPI"))
mydata2$position_9 <- factor(mydata2$position_9, levels = c("","CV-gap", "NonSci", "NonRes", "IndR", "AcPD", "AcOt", "AcPI"))
mydata2$position_13 <- factor(mydata2$position_13, levels = c("","CV-gap", "NonSci", "NonRes", "IndR", "AcPD", "AcOt", "AcPI"))
mydata2$position_17 <- factor(mydata2$position_17, levels = c("","CV-gap", "NonSci", "NonRes", "IndR", "AcPD", "AcOt", "AcPI"))
predoc <- mydata2[ which(mydata2$type_pre_postdoc=='predoc'),]
postdoc <- mydata2[ which(mydata2$type_pre_postdoc=='postdoc'),]

##change colours###
mycols_pos1<-c("#D0D0CE","#E2E868","#F4C61F","#F49E17","#D41645","#734595","#3B6FB6")
names(mycols_pos1)=c('CV-gap','NonSci','NonRes','IndR','AcPD','AcOt','AcPI')
mycols<-c("white","#D0D0CE","#E2E868","#F4C61F","#F49E17","#D41645","#734595","#3B6FB6")
names(mycols)=c('','CV-gap','NonSci','NonRes','IndR','AcPD','AcOt', 'AcPI')
```

### Positions after five years (Figure 2A)

```{r fig2A, fig.width=8, fig.height=4, dpi=200}
#postdoc_5
postdoc_cohort <- postdoc %>%
  dplyr::group_by(cohort, position_5) %>%
  dplyr::summarise(count = n())

bp <- ggplot(postdoc_cohort[which(postdoc_cohort$position_5!=""),], 
             aes(fill=position_5, y=count, x=cohort),col=mycols) + 
  geom_bar( stat="identity", position="fill") + 
  scale_fill_manual(values=mycols) + 
  labs(x = "cohort", y="Proportion in each role") + 
  ggtitle("Postdocs, 
5 years after EMBL")+ theme_minimal(base_size = 16) + theme(axis.title=element_text(size=12)) + theme(plot.title = element_text(size=16), axis.text.x = element_text(size = 12,angle=45, hjust=1))+ 
  theme(axis.title=element_text(size=12)) + theme(plot.title=element_text(size=12))+
  theme(legend.key.height = unit(1, 'lines'),legend.key.width = unit(0.4, 'lines'),legend.title = element_text(size=12), legend.text = element_text(size=12)) 
  

#predoc_5
predoc_cohort <- predoc %>%
  dplyr::group_by(cohort, position_5) %>%
  dplyr::summarise(count = n())

b <- ggplot(predoc_cohort[which(predoc_cohort$position_5!=""),], 
            aes(fill=position_5, y=count, x=cohort),col=mycols) + 
  geom_bar(stat="identity", position="fill") + 
  scale_fill_manual(values=mycols) + labs(x = "cohort", y="Proportion in each role") +
  ggtitle("PhD students, 
5 years after EMBL")+ theme_minimal(base_size = 16) + theme(axis.title=element_text(size=12)) + theme(plot.title = element_text(size=12), axis.text.x = element_text(size = 12,angle=45, hjust=1))+ 
  theme(axis.title=element_text(size=12)) + theme(plot.title=element_text(size=12))+
  theme(legend.key.height = unit(1, 'lines'),legend.key.width = unit(0.4, 'lines'),legend.title = element_text(size=12), legend.text = element_text(size=12)) 

#extract legend
g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

for_legend <-bp+theme(legend.position="right")
mylegend<-g_legend(for_legend)


grid.arrange(arrangeGrob(b + theme(legend.position="none"),
                               bp + theme(legend.position="none"),
                               mylegend, 
                               nrow=1,widths=c(5,5,2)))
```

### Other time points (Supplementary Figure 2)
```{r supp2, fig.height=20, fig.width=16, dpi=200}
#postdoc_1
postdoc_cohort <- postdoc %>%
  dplyr::group_by(cohort, position_1) %>%
  dplyr::summarise(count = n())

ap <- ggplot(postdoc_cohort[which(postdoc_cohort$position_1!=""),], 
             aes(fill=position_1, y=count, x=cohort),col=mycols_pos1) + 
  geom_bar( stat="identity", position="fill") + 
  scale_fill_manual(values=mycols_pos1) + 
  labs(x = "cohort", y="Proportion in each role") + 
  ggtitle("Postdocs, 1 year after EMBL") + 
  theme_minimal(base_size = 20) + theme(plot.title = element_text(size=16))

#postdoc_5
postdoc_cohort <- postdoc %>%
  dplyr::group_by(cohort, position_5) %>%
  dplyr::summarise(count = n())

bp <- ggplot(postdoc_cohort[which(postdoc_cohort$position_5!=""),], 
             aes(fill=position_5, y=count, x=cohort),col=mycols) + 
  geom_bar( stat="identity", position="fill") + 
  scale_fill_manual(values=mycols) + 
  labs(x = "cohort", y="Proportion in each role") + 
  ggtitle("Postdocs, 5 years after EMBL") + 
  theme_minimal(base_size = 20) + theme(plot.title = element_text(size=16))

#postdoc_9
postdoc_cohort <- postdoc %>%
  dplyr::group_by(cohort, position_9) %>%
  dplyr::summarise(count = n())

cp<-ggplot(postdoc_cohort[which(postdoc_cohort$position_9!=""),], 
           aes(fill=position_9, y=count, x=cohort),col=mycols) + 
  geom_bar(stat="identity", position="fill") + 
  scale_fill_manual(values=mycols) + 
  labs(x = "cohort", y="Proportion in each role") + 
  ggtitle("Postdocs, 9 years after EMBL") + 
  theme_minimal(base_size = 20) + theme(plot.title = element_text(size=16))

#postdoc_13
postdoc_cohort <- postdoc %>%
  dplyr::group_by(cohort, position_13) %>%
  dplyr::summarise(count = n())

dp <- ggplot(postdoc_cohort[which(postdoc_cohort$position_13!=""),], 
             aes(fill=position_13, y=count, x=cohort),col=mycols) + 
  geom_bar( stat="identity", position="fill") + 
  scale_fill_manual(values=mycols) + 
  labs(x = "cohort", y="Proportion in each role") + 
  ggtitle("Postdocs, 13 years after EMBL") + 
  theme_minimal(base_size = 20)+ theme(plot.title = element_text(size=16))

#postdoc_17
postdoc_cohort <- postdoc %>%
  dplyr::group_by(cohort, position_17) %>%
  dplyr::summarise(count = n())

ep <- ggplot(postdoc_cohort[which(postdoc_cohort$position_17!=""),], 
            aes(fill=position_17, y=count, x=cohort),col=mycols) + 
  geom_bar( stat="identity", position="fill") + scale_fill_manual(values=c(mycols)) +
  labs(x = "cohort", y="Proportion in each role") + 
  ggtitle("Postdocs, 17 years after EMBL")+ 
  theme_minimal(base_size = 20) + theme(plot.title = element_text(size=16))



#predoc_1
predoc_cohort <- predoc %>%
  dplyr::group_by(cohort, position_1) %>%
  dplyr::summarise(count = n())

a <- ggplot(predoc_cohort[which(predoc_cohort$position_1!=""),], 
            aes(fill=position_1, y=count, x=cohort),col=mycols_pos1) + 
  geom_bar(stat="identity", position="fill") + scale_fill_manual(values=mycols_pos1) + 
  labs(x = "cohort", y="Proportion in each role") + ggtitle("PhD students, 1 year after EMBL") + 
  theme_minimal(base_size = 20) + theme(plot.title = element_text(size=16))

#predoc_5
predoc_cohort <- predoc %>%
  dplyr::group_by(cohort, position_5) %>%
  dplyr::summarise(count = n())

b <- ggplot(predoc_cohort[which(predoc_cohort$position_5!=""),], 
            aes(fill=position_5, y=count, x=cohort),col=mycols) + 
  geom_bar(stat="identity", position="fill") + scale_fill_manual(values=mycols) + 
  labs(x = "cohort", y="Proportion in each role") + ggtitle("PhD students, 5 years after EMBL") + 
  theme_minimal(base_size = 20) + theme(plot.title = element_text(size=16))

#predoc_9
predoc_cohort <- predoc %>%
  dplyr::group_by(cohort, position_9) %>%
  dplyr::summarise(count = n())

c <- ggplot(predoc_cohort[which(predoc_cohort$position_9!=""),], 
            aes(fill=position_9, y=count, x=cohort),col=mycols) + 
  geom_bar( stat="identity", position="fill") + scale_fill_manual(values=mycols) + 
  labs(x = "cohort", y="Proportion in each role") + 
  ggtitle("PhD students, 9 years after EMBL") + 
  theme_minimal(base_size = 20) + theme(plot.title = element_text(size=16))

#predoc_13
predoc_cohort <- predoc %>%
  dplyr::group_by(cohort, position_13) %>%
  dplyr::summarise(count = n())

d <- ggplot(predoc_cohort[which(predoc_cohort$position_13!=""),], 
            aes(fill=position_13, y=count, x=cohort),col=mycols) + 
  geom_bar(stat="identity", position="fill") + scale_fill_manual(values=mycols) + 
  labs(x = "cohort", y="Proportion in each role") + ggtitle("PhD students, 13 years after EMBL") + 
  theme_minimal(base_size = 20) + theme(plot.title = element_text(size=16))

#predoc_17
predoc_cohort <- predoc %>%
  dplyr::group_by(cohort, position_17) %>%
  dplyr::summarise(count = n())

e <- ggplot(predoc_cohort[which(predoc_cohort$position_17!=""),], 
            aes(fill=position_17, y=count, x=cohort),col=mycols) + 
  geom_bar( stat="identity", position="fill") + scale_fill_manual(values=c(mycols)) + 
  labs(x = "cohort", y="Proportion in each role") + ggtitle("PhD students, 17 years after EMBL") + 
  theme_minimal(base_size = 20) + theme(plot.title = element_text(size=16))



#combine the figures
grid.arrange(arrangeGrob(a + theme(legend.position="none"),
                         ap + theme(legend.position="none"),
                         b + theme(legend.position="none"),
                         bp + theme(legend.position="none"),
                         c + theme(legend.position="none"),
                         cp + theme(legend.position="none"),
                         d + theme(legend.position="none"),
                         dp + theme(legend.position="none"),
                         e + theme(legend.position="none"),
                         ep + theme(legend.position="none") , nrow=5,ncol=2))
```

## Compare with other cohorts (Figure 2B and Supplementary Figure 3)

### Data preparation for plotting
```{r}
#load data file with values
otherCohort = read_csv("data/in_year_all.csv")
my_col <- c("grey40","#3B6FB6")
```

```{r fig2b, fig.width=6, fig.height=5, dpi=200}

  mydata <- melt(otherCohort[which (otherCohort$institution=="Stanford University (SU) Bioscience | 2018 dataset."),], id=c("institution","Cohorts","Cohort_short","ratio", "ratio_text"))
  mydata$variable <- factor(mydata$variable , levels = c("EMBL", "NorthAmerica"))
  
  levels(mydata$variable) <- c("EMBL", "Stanford University (SU) Bioscience | 2018 dataset.")
    plot_b<-ggplot(mydata, aes(x=Cohorts, y=value, fill=variable)) +geom_bar(position="dodge",stat="identity") +theme_bw(base_size = 20)+ylim(0,0.5)  + theme(legend.position="top")+
  guides(fill = guide_legend(nrow = 2, byrow = TRUE, title.position="top")) +labs(x = "cohort", y="Alumni in PI-like roles
(proportion)")+ scale_fill_manual(values=c("#3B6FB6","grey40"),name = "Institution:")+  theme(legend.key.height = unit(0.6, 'lines'),legend.title = element_text(size=13), legend.text = element_text(size=13))
  
  plot_d<-ggplot(mydata[ which(mydata$variable=='EMBL'),], aes(x=Cohort_short, y=ratio)) +geom_point() +
    theme_bw(base_size = 12)+geom_hline(yintercept = 1, linetype=3)+
 theme( axis.text.x = element_text(size = 10),plot.background = element_rect(fill = "grey90"))+labs(x = "cohort", y="EMBL / SU")+ scale_y_continuous(breaks=seq(0,2,1))+ coord_cartesian(ylim=c(-.2, 2.2))
  
  mydata_plot <-  ggdraw() +
    draw_plot(plot_b) +
    draw_plot(plot_d, x = 0.72, y = .35, width= 0.25, height = .35)
  
  mydata_plot


```


```{r supp3, fig.width=16, fig.height=20, dpi=400}
institions<-sort(unique(otherCohort$institution[otherCohort$institution!="Stanford University (SU) Bioscience | 2018 dataset."]))

pList <- lapply(institions, function(i) {
  
  mydata <- melt(otherCohort[which (otherCohort$institution==i),], id=c("institution","Cohorts","Cohort_short","ratio","ratio_text"))
  mydata$variable <- factor(mydata$variable , levels = c("EMBL", "NorthAmerica"))
  yaxistext<-unique(mydata$ratio_text)
  levels(mydata$variable) <- c("EMBL", i)
    plot_b<-ggplot(mydata, aes(x=Cohorts, y=value, fill=variable)) +geom_bar(position="dodge",stat="identity") +theme_bw(base_size = 22)+ylim(0,0.5) +
    theme(legend.position="top")+
  guides(fill = guide_legend(nrow = 2, byrow = TRUE, title.position="left")) +labs(x = "cohort", y="Alumni in PI-like roles
(proportion)")+ scale_fill_manual(values=c("#3B6FB6","grey40"),name = "Institution:")
  
  plot_d<-ggplot(mydata[ which(mydata$variable=='EMBL'),], aes(x=Cohort_short, y=ratio)) +geom_point() +
    theme_bw(base_size = 16)+geom_hline(yintercept = 1, linetype=3)+
 theme( axis.text.x = element_text(size = 14),plot.background = element_rect(fill = "grey90"))+labs(x = "cohort", y=yaxistext)+ scale_y_continuous(breaks=seq(0,2,1))+ coord_cartesian(ylim=c(-.2, 2.2))
  
  width_count<-0.05+(0.075*length(unique(mydata$Cohort_short)))
  
  mydata_plot <-  ggdraw() +
    draw_plot(plot_b) +
    draw_plot(plot_d, x = 0.7, y = .4, width = width_count, height = .3)
  
  mydata_plot
})

plot_grid(plotlist = pList, ncol=2)

```


## Kaplan-Meier plots of time to position, stratified by cohorts

### AcPI
```{r km_AcPI_cohort,fig.width=6, fig.height=4, dpi=200}
corhortOut <- plotCohort(careerData, "AcPI", showTable = FALSE)
corhortOut$plot
tab1_co <- corhortOut$table %>% mutate(Position = "AcPI")
```

### AcOt
```{r km_AcOt_cohort,fig.width=6, fig.height=4, dpi=200}
corhortOut <- plotCohort(careerData, "AcOt", showTable = FALSE)
corhortOut$plot
tab2_co <- corhortOut$table %>% mutate(Position = "AcOt")
```

### IndR
```{r km_IndR_cohort,fig.width=6, fig.height=4, dpi=200}
corhortOut <- plotCohort(careerData, "IndR", showTable = FALSE)
corhortOut$plot
tab3_co <- corhortOut$table %>% mutate(Position = "IndR")
```


### NonRes
```{r km_NonRes_cohort, fig.width=6, fig.height=4, dpi=200}
corhortOut <- plotCohort(careerData, "NonRes", showTable = FALSE)
corhortOut$plot
tab4_co <- corhortOut$table %>% mutate(Position = "NonRes")
```

### NonSci
```{r km_NonSci_cohort, fig.width=6, fig.height=4, dpi=200}
corhortOut <- plotCohort(careerData, "NonSci", showTable = FALSE)
corhortOut$plot
tab5_co <- corhortOut$table %>% mutate(Position = "NonSci")
```

## Simplified career paths for alumni who have held different types of role (Supplementry Figure 1)

```{r}
#configuration of python environment
use_condaenv("r-reticulate")
```
### Color code for the Sankey diagram
```{r}
colList <- list(NonSci = "#E2E868", NonRes = "#F4C61F",IndR = "#F49E17",AcPD = "#D41645", AcPhD = "#707372", AcOt = "#734595",  AcPI = "#3B6FB6")
             
dummyTab <- data.frame(var = names(colList), val = 1)
g <- ggplot(dummyTab, aes(x=var, y=val, fill = var )) + geom_bar(stat="identity") +
  scale_fill_manual(values = colList, name = "role")
legendPlot <- cowplot::get_legend(g)
plot_grid(legendPlot)
```

Function for producing Sankey diagram for career path 
```{r}
```


```{r}
```


```{r}
plotSankey <- function(tempdata, role) {
  
  rolePrev <- list(was_GL = "new_AcFac_previous",
                   CV_wasNonSci = "new_NonSci_previous",
                   CV_wasIndR = "new_IndR_previous", 
                   CV_wasNonR = "new_SciR_previous",
                   CV_wasAcOt = "new_AcOt_previous")
  
  #first filter for alumni with full CV who became PIs
  Type_data <- tempdata[which(tempdata$completeness=='Complete CV'),]
  Type_data <- Type_data[which(Type_data[[role]]=='y'),]

  #create summary tables for previous and current roles and swap names for nodes
  Type_count <- Type_data  %>% dplyr::group_by_at(rolePrev[[role]]) %>% dplyr::summarize(count = n())
  
  Type_count <- add_column(Type_count, target = 7, .after = 1)
  Type_count<-as.data.frame(Type_count)
  names(Type_count)[1]<- 'source'
  names(Type_count)[3]<- "value"
  Type_count$source = gsub("AcFac",0,Type_count$source)
  Type_count$source = gsub("AcOt",1,Type_count$source)
  Type_count$source = gsub("AcPD",2,Type_count$source)
  Type_count$source = gsub("AcPhD",3,Type_count$source)
  Type_count$source = gsub("IndR",4,Type_count$source)
  Type_count$source = gsub("NonRes",5,Type_count$source)
  Type_count$source = gsub("NonSci",6,Type_count$source)

  Type_now <- Type_data  %>%
    dplyr::group_by(timepointLast) %>%
    dplyr::summarize(count = n())
  Type_now <- add_column(Type_now, source = 7, .before = 1)
  Type_now<-as.data.frame(Type_now)
  names(Type_now)[2]<- 'target'
  names(Type_now)[3]<- "value"
  Type_now$target = gsub("AcFac",14,Type_now$target)
  Type_now$target = gsub("AcOt",15,Type_now$target)
  Type_now$target = gsub("AcPD",16,Type_now$target)
  Type_now$target = gsub("AcPhD",17,Type_now$target)
  Type_now$target = gsub("IndR",18,Type_now$target)
  Type_now$target = gsub("NonRes",19,Type_now$target)
  Type_now$target = gsub("NonSci",20,Type_now$target)

  #merge and plot
  links <- rbind(Type_count, Type_now)
  
  p <- plot_ly(
    type = "sankey",
    orientation = "h",
    
    node = list(
      label = c("","", "", "","","","",
                "","","","","","","",
                "", "","","","","",""),
      
      color = c(colList$AcPI, colList$AcOt, colList$AcPD, colList$AcPhD, colList$IndR, colList$NonRes, colList$NonSci,
                '#000000', '#000000', '#000000', '#000000', '#000000', '#000000', '#000000',
                colList$AcPI, colList$AcOt, colList$AcPD, colList$AcPhD, colList$IndR, colList$NonRes, colList$NonSci),
      pad = 25,
      thickness = 75,
      line = list(
        color = "black",
        width = 0
      )
    ),
    
    link =links
    )
  p
}
```


### AcPI
```{r}
plotSankey(originalData, "was_GL")
```

### NonSci
```{r}
plotSankey(originalData, "CV_wasNonSci")
```

### IndR
```{r}
plotSankey(originalData, "CV_wasIndR")
```

### NonRes
```{r}
plotSankey(originalData, "CV_wasNonR")
```

### AcOt
```{r}
plotSankey(originalData, "CV_wasAcOt")
```

**The text labels in this supplementary figure are added manually, and the central bar re-coloured to match the key (e.g. blue for the AcPI plot)**




Box-plot with overlaid dot-plot showing the distribution in calendar years between PhD conferral and start of first PI role for PhD alumni who defended in 1997-2012 who became a principal investigator within 9-years, and for whom we have a detailed career path (n=157).   
```{r predocGL, fig.height=4, fig.width=6, dpi=200}
mydatatime <- as.data.frame(tempdata) 
mydatatime <- mydatatime[ which(mydatatime$completeness=="Complete CV" ), ]
predoc_data <- mydatatime[ which(mydatatime$type_pre_postdoc=='predoc'), ]
predoc_to2012 <- predoc_data[ which(predoc_data$to_year<2013), ]
predoc_to2012_exit_GL <-predoc_to2012[ which(predoc_to2012$PhDtoGL<10& predoc_to2012$completeness=="Complete CV" ), ]
predoc_to2012_exit_GL$cohort <- as.factor(predoc_to2012_exit_GL$cohort)

stat <- t.test(predoc_to2012_exit_GL$PhDtoGL~predoc_to2012_exit_GL$cohort, var.equal = FALSE)[["p.value"]]
f10 <- -log10(stat)
fstat <- round (stat, f10 +3)
fstat[stat<2.2e-16] <- "<2.2e-16"
fstat <- paste("p=", fstat, " ", if (stat<0.05) {"*"},if (stat<0.001) {"*"},if (stat<0.0001) {"*"}, sep="")
fstat <- toString(fstat)



p <- ggboxplot(predoc_to2012_exit_GL, x="cohort", y="PhDtoGL", fill='#ffffff', color="grey58")
p <- p + geom_dotplot(aes(predoc_to2012_exit_GL$cohort, predoc_to2012_exit_GL$PhDtoGL), 
                      binaxis='y', stackdir='center',stackratio=1.0, dotsize=0.25) + 
  stat_summary(fun.y = "mean", geom = "point", color = "RED", shape=4, size=10)  + 
  theme_classic(base_size = 14) + geom_hline(yintercept = 9,linetype="dashed")+ scale_y_continuous(breaks=seq(0,9,3))+coord_cartesian(ylim=c(0, 10))+ggtitle(fstat)+theme(plot.title = element_text(size=12))
p
```


Box-plot with overlaid dot-plot showing the distribution in calendar years between completion of an EMBL postdoc and start of first PI role for postdoc alumni who completed their postdoc in 1997-2012 who became a principal investigator within 9-years, and for whom we have a detailed career path (n=182).  
```{r postdocphdgl, fig.height=4, fig.width=6, dpi=200}
# postdoc by cohort
postdoc_data <- mydatatime[ which(mydatatime$type_pre_postdoc=='postdoc'), ]
data_sum <- postdoc_data %>% dplyr::group_by(cohort,PhDtoGL) %>% dplyr::summarise(n = n())
postdoc_to2012 <- postdoc_data[ which(postdoc_data$to_year<2013), ]
postdoc_to2012_exit_GL <- postdoc_to2012[ which(postdoc_to2012$PhDtoGL<10 & postdoc_to2012$completeness=="Complete CV" ), ]
postdoc_to2012_exit_GL$cohort <- as.factor(postdoc_to2012_exit_GL$cohort)
postdoc_to2012_exit_GLEMBL <- postdoc_to2012[ which(postdoc_to2012$EMBLtoPI<10 & postdoc_to2012$EMBLtoPI>-1  & postdoc_to2012$completeness=="Complete CV" ), ]
postdoc_to2012_exit_GLEMBL$cohort <- as.factor(postdoc_to2012_exit_GLEMBL$cohort)

stat <- t.test(postdoc_to2012_exit_GLEMBL$EMBLtoPI~postdoc_to2012_exit_GLEMBL$cohort, var.equal = FALSE)[["p.value"]]
f10 <- -log10(stat)
fstat <- round (stat, f10 +3)
fstat[stat<2.2e-16] <- "<2.2e-16"
fstat <- paste("p=", fstat, " ", if (stat<0.05) {"*"},if (stat<0.001) {"*"},if (stat<0.0001) {"*"}, sep="")
fstat <- toString(fstat)

p <- ggboxplot(postdoc_to2012_exit_GLEMBL, x="cohort", y="EMBLtoPI", fill='#ffffff', color="grey58")
p <- p + geom_dotplot(aes(postdoc_to2012_exit_GLEMBL$cohort, postdoc_to2012_exit_GLEMBL$EMBLtoPI), 
                      binaxis='y', stackdir='center',stackratio=1.0, dotsize=0.25) +
  stat_summary(fun.y = "mean", geom = "point", color = "RED", shape=4, size=10)  + 
  theme_classic(base_size = 14) + geom_hline(yintercept = 9,linetype="dashed")+ scale_y_continuous(breaks=seq(0,9,3))+coord_cartesian(ylim=c(0, 10))+ggtitle(fstat)+theme(plot.title = element_text(size=12))
p
```

 
Same as above but for postdoc alumni for whom we were able to identify a PhD conferral year
```{r figpostdocGL,  fig.height=4, fig.width=6, dpi=200}
stat <- t.test(postdoc_to2012_exit_GL$PhDtoGL~postdoc_to2012_exit_GL$cohort, var.equal = FALSE)[["p.value"]]

f10 <- -log10(stat)
fstat <- round (stat, f10 +3)
fstat[stat<2.2e-16] <- "<2.2e-16"
fstat <- paste("p=", fstat, " ", if (stat<0.05) {"*"},if (stat<0.001) {"*"},if (stat<0.0001) {"*"}, sep="")
fstat <- toString(fstat)

p <- ggboxplot(postdoc_to2012_exit_GL, x="cohort", y="PhDtoGL", fill='#ffffff', color="grey58")
p <- p + geom_dotplot(aes(postdoc_to2012_exit_GL$cohort, postdoc_to2012_exit_GL$PhDtoGL), 
                      binaxis='y', stackdir='center',stackratio=1.0, dotsize=0.25)+
  stat_summary(fun.y = "mean", geom = "point", color = "RED", shape=4, size=10)  + labs(x="Postdoc cohort", y = "Calendar years, PhD to first PI role") + 
  theme_classic(base_size = 14)+geom_hline(yintercept = 9,linetype="dashed")+ scale_y_continuous(breaks=seq(0,9,3))+coord_cartesian(ylim=c(0, 10))+ggtitle(fstat)+theme(plot.title = element_text(size=12))
p


```


# Section 3: Predictors of academic PI positions 

## Figure 3A

Histograms of the number of first author EMBL publications per ECR for alumni who have not (yet) become a PI (PI) and those who became a PI.  

Plot
```{r fig3a, fig.height=4, fig.width=5, dpi=200}
mydata <- as.data.frame(tempdata) 
mydata$was_GL = replace_na(mydata$was_GL,"n")

#calculate mean by group
df_mean <- mydata %>% group_by(was_GL) %>% summarise(meant = mean(pubs_FIRST_ra_only_TOTAL))
upper_lim<-quantile(mydata$pubs_FIRST_ra_only_TOTAL, 0.975)
p <- ggplot(mydata, aes(x=pubs_FIRST_ra_only_TOTAL)) + 
  geom_histogram(binwidth=1, colour="black", fill="grey80") + facet_grid(was_GL ~ .) +
  theme(strip.text.y = element_text(angle = 0)) + 
  labs(x = "# first-author research articles", y="number of alumni") +
  theme_bw(base_size = 16)

p<- p +  geom_vline(data=filter(df_mean, was_GL=="n"), 
                    aes(xintercept=meant), colour="red",linetype="dashed") + 
  geom_vline(data=filter(df_mean, was_GL=="y"), aes(xintercept=meant), colour="red",linetype="dashed")

p <- p +  geom_text(data=filter(df_mean, was_GL=="n"), aes(label = round(meant,1), x = meant, y = 400),  
                    size=6,colour="red", angle=-90, vjust = -1) +  
  geom_text(data=filter(df_mean, was_GL=="y"), aes(label = round(meant,1), x = meant, y = 400),
            size=6,colour="red", angle=-90, vjust = -1) + xlim(-0.5,upper_lim+0.5)

#do t.test and extract p-value
stat <- t.test(mydata$pubs_FIRST_ra_only_TOTAL~mydata$was_GL, var.equal = FALSE)[["p.value"]]
f10 <- -log10(stat)
fstat <- round (stat, f10 +3)
fstat[stat<2.2e-16] <- "<2.2e-16"
fstat <- paste("p=", fstat, " ", if (stat<0.05) {"*"},if (stat<0.001) {"*"},if (stat<0.0001) {"*"}, sep="")
fstat <- toString(fstat)

#plot
p <- p + ggtitle(fstat)+theme(plot.title = element_text(size=12), axis.text=element_text(size=12))
p

```

Summary statistics
```{r}
summary_stats <-group_by(mydata, was_GL) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(pubs_FIRST_ra_only_TOTAL, na.rm = TRUE),
    sd = sd(pubs_FIRST_ra_only_TOTAL, na.rm = TRUE)
  )
```

## Figure 3B   

Histograms of the natural logarithm of the highest category normalized citation of first author EMBL publications per ECR publishing at least one first author paper linked to EMBL and with a CNCI value in Clarivate InCites’ database) for alumni who have not (yet) become a PI (PI) and those who became a PI. 

```{r fig3b,  fig.height=4, fig.width=5, dpi=200}
df_mean <- mydata %>% group_by(was_GL) %>% summarise(meant = mean(pubs_FIRST_RA_CNCI_max, na.rm=TRUE))
df_mean

p <- ggplot(mydata, aes(x=log(pubs_FIRST_RA_CNCI_max))) + 
  geom_histogram(binwidth=1, colour="black", fill="grey80") +facet_grid(was_GL ~ .) +
  theme(strip.text.y = element_text(angle = 0)) + 
  labs(x = "log(max CNCI, first-author research articles)", y="number of alumni")+
  theme_bw(base_size = 16)

p<- p +  
  geom_vline(data=filter(df_mean, was_GL=="n"), aes(xintercept=log(meant)), colour="red",linetype="dashed") + 
  geom_vline(data=filter(df_mean, was_GL=="y"), aes(xintercept=log(meant)), colour="red",linetype="dashed")

p<- p +  geom_vline(aes(xintercept=log(1)), colour="black") 
p<- p +  geom_text(data=filter(df_mean, was_GL=="n"), aes(label = paste0("log(",round(meant,1),")"), x = log(meant), y = 250),  
                  size=6,colour="red", angle=-90, vjust = -1) +  
  geom_text(data=filter(df_mean, was_GL=="y"), aes(label = paste0("log(",round(meant,1),")"), x = log(meant), y = 250),
            size=6,colour="red", angle=-90, vjust = -1) 


#do t.test and extract p-value
stat<-t.test(mydata$pubs_FIRST_RA_CNCI_max~mydata$was_GL, var.equal = FALSE)[["p.value"]]
f10 <- -log10(stat)
fstat <- round (stat, f10 +3)
fstat[stat<2.2e-16] <- "<2.2e-16"
fstat <- paste("p=", fstat, " ", if (stat<0.05) {"*"},if (stat<0.001) {"*"},if (stat<0.0001) {"*"}, sep="")
fstat <- toString(fstat)

#plot
p<-p + ggtitle(fstat)+theme(plot.title = element_text(size=12), axis.text=element_text(size=12))
p
```

Summary statistics
```{r}
summary_stats <-group_by(mydata, was_GL) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(pubs_FIRST_RA_CNCI_max, na.rm = TRUE),
    sd = sd(pubs_FIRST_RA_CNCI_max, na.rm = TRUE)
  )

summary_stats

```

## Publication versus PI (Figure 3D)

### Stratified by PhD and postdocs

Preprocessing data
```{r}

#get survival table
survT <- processSurvivalTable(careerData, "AcPI","EMBL")

testTab <- survT %>% 
  left_join(select(careerData, unique_ID, type_pre_postdoc, cohort, pubs_FIRST_ra_only_TOTAL), by = "unique_ID") %>%
  filter(!is.na(timeToPos), !is.na(pubs_FIRST_ra_only_TOTAL), !is.na(type_pre_postdoc)) %>%
  mutate(cohort = factor(cohort),
         type_pre_postdoc = ifelse(type_pre_postdoc == "predoc","PhD alumni", "Postdoc alumni")) %>%
  mutate(numPub = ifelse(pubs_FIRST_ra_only_TOTAL > 1, "2+", pubs_FIRST_ra_only_TOTAL)) %>%
  mutate(type_pre_postdoc = factor(type_pre_postdoc, levels = c("PhD alumni","Postdoc alumni")),
         numPub = factor(numPub, levels = c("0", "1", "2+")))
  
print("Total subject number stratified by pre/post-doc")
print(table(testTab$type_pre_postdoc))
print("Event number stratified by pre/post-doc")
print(table(filter(testTab, ifPos)$type_pre_postdoc))
```  

Plot KM curves stratified by pre-postdoc

```{r km_AcPI_PubNum, fig.height=4, fig.width=8, dpi=200}

plotOut <- km(testTab, "numPub",  maxTime = 25, titlePlot = "",
       titleLegend = "Number of\nfirst-author\npublications",
       xlab = sprintf("Time after %s (years)", "EMBL"),
       ylab = paste0("Probability of being found as ",  "AcPI")) +
  theme(legend.position = "right")



plotOut
```


## Quantify the power of predictors using Harrells’ C-Index (Figure 3C)

### AcPI
```{r HC_both_AcPI, fig.height=5, fig.width=6, cache=TRUE, dpi=200}
p <- plotHarrelsC(careerData, "AcPI", preOrPost = "both")
p$plot
tabOut <- p$table

pubOut <- plotPublication(careerData, "AcPI", showTable = FALSE)
tab1 <- pubOut$table %>% mutate(Position = "AcPI")
```

# Section 4: predictors for non-PI positions

## Number of publications as a predictor, stratified by PhD and postdocs (Figure 4C and Supplementary Figure 5)

### AcOt
```{r km_AcOt_numPub,fig.width=6, fig.height=4, dpi=200}
pubOut <- plotPublication(careerData, "AcOt", showTable = FALSE)
pubOut$plot
tab2 <- pubOut$table %>% mutate(Position = "AcOt")
```

### IndR
```{r km_IndR_numPub,fig.width=6, fig.height=4, dpi=200}
pubOut <- plotPublication(careerData, "IndR", showTable = FALSE)
pubOut$plot
tab3 <- pubOut$table %>% mutate(Position = "IndR")
```


### NonRes
```{r km_NonRes_numPub, fig.width=6, fig.height=4, dpi=200}
pubOut <- plotPublication(careerData, "NonRes", showTable = FALSE)
pubOut$plot
tab4 <- pubOut$table %>% mutate(Position = "NonRes")
```

### NonSci
```{r km_NonSci_numPub, fig.width=6, fig.height=4, dpi=200}
putOut <- plotPublication(careerData, "NonSci", showTable = FALSE)
putOut$plot
tab5 <- putOut$table %>% mutate(Position = "NonSci")
```

## Quantify the power of predictors using Harrells’ C-Index (Figure 4A,B,D,E)

### AcOt
```{r HC_both_AcOt, fig.height=5, fig.width=6, cache=TRUE, dpi=200}
p <- plotHarrelsC(careerData, "AcOt", preOrPost = "both")
p$plot
tabOut <- bind_rows(tabOut, p$table)
```

### IndR
```{r HC_both_IndR, fig.height=5, fig.width=6, cache=TRUE, dpi=200}
p <- plotHarrelsC(careerData, "IndR", preOrPost = "both")
p$plot
tabOut <- bind_rows(tabOut, p$table)
```

### NonRes
```{r HC_both_NonRes, fig.height=5, fig.width=6, cache=TRUE, dpi=200}
p <- plotHarrelsC(careerData, "NonRes", preOrPost = "both")
p$plot
tabOut <- bind_rows(tabOut, p$table)
```

### NonSci
```{r HC_both_NonSci, fig.height=5, fig.width=6, cache=TRUE, dpi=200}
p <- plotHarrelsC(careerData, "NonSci", preOrPost = "both")
p$plot
tabOut <- bind_rows(tabOut, p$table)
```

# Section 5: Publications are increasingly collaborative


## Figure 5B and 5C

## Figure 5B and 5C

```{r fig5B, fig.height=4, fig.width=5, dpi=200}
#Cohort vs ra articles - boxplot 
#do anova on all data, extract p-value

stat<-aov(tempdata$pubs_RAonly_TOTAL~tempdata$cohort)
stat<- summary(stat)[[1]][["Pr(>F)"]][[1]]
f10 <- -log10(stat)
fstat <-round (stat, f10 +3)
fstat<-paste("ANOVA p=", fstat, " ", if (stat<0.05) {"*"},if (stat<0.001) {"*"},if (stat<0.0001) {"*"}, sep="")
fstat <- toString(fstat)

levels(tempdata$cohort)
#plot, excluding long-tail outliers
upper_lim<-quantile(tempdata$pubs_RAonly_TOTAL, 0.975)
p1 <- ggplot(tempdata, aes(x=cohort, y=pubs_RAonly_TOTAL)) + 
  geom_boxplot(lwd=0.5,color="grey40", fill="grey100",  width=0.5)+
  stat_summary(fun.y = "mean", geom = "point",shape=4, color="red",size=4) + 
  theme_classic(base_size = 14) +
  theme(panel.grid.major=element_line(colour="grey90")) +
  labs(x = "cohort", y="# research articles")
p1<-p1+ ggtitle(fstat)+ylim(0,upper_lim)+ theme(plot.title=element_text(size=12)) 
p1
upper_lim<-quantile(tempdata$pubs_RAonly_TOTAL, 0.975)

##Cohort vs first.author ra articles - boxplot 
#do anova on all data, extract p-value
stat<-aov(tempdata$pubs_FIRST_ra_only_TOTAL~tempdata$cohort)
stat<- summary(stat)[[1]][["Pr(>F)"]][[1]]
f10 <- -log10(stat)
fstat <-round (stat, f10 +3)
fstat<-paste("ANOVA p=", fstat, " ", if (stat<0.05) {"*"},if (stat<0.001) {"*"},if (stat<0.0001) {"*"}, sep="")
fstat <- toString(fstat)
#plot, excluding long-tail outliers
upper_lim<-quantile(tempdata$pubs_FIRST_ra_only_TOTAL, 0.975)
p2 <- ggplot(tempdata, aes(x=cohort, y=pubs_FIRST_ra_only_TOTAL)) + 
  geom_boxplot(lwd=0.5,color="grey40", fill="grey100",  width=0.5)+
  stat_summary(fun.y = "mean", geom = "point",shape=4, color="red",size=4) + 
  theme_classic(base_size = 14) +
  theme(panel.grid.major=element_line(colour="grey90")) +
  labs(x = "cohort", y="# first-author research articles")
p2<-p2+ ggtitle(fstat)+ylim(0,upper_lim)+ theme(plot.title=element_text(size=12)) 
p2

```


## Figure 5A and 5D
```{r fig5ab, fig.height=4, fig.width=5, dpi=200}
pubdata <- read_csv("data/WOSpapers_unique.csv") %>%
  rename_with(make.names)

pubdata_art <- pubdata[ which(pubdata$articletype=='Article'), ]
pubdata_oth <- pubdata[ which(pubdata$articletype=='Other'), ]

a<- dplyr::group_by(pubdata_art, Publication.Date) %>%
  dplyr::summarise(
    count = n(),
    wmean = winsor.means(count_authors, na.rm = TRUE),
    mean = mean(count_authors, na.rm = TRUE),
    median = median(count_authors, na.rm = TRUE),
    sd = sd(count_authors, na.rm = TRUE)
  )

# Create Line Chart
p <- ggplot(a, aes(Publication.Date, y=wmean)) + geom_point(color = "black") + 
  ylim(0,12)+ xlim(1995,2020) + 
  labs(x="Year", y = "Winsorized mean,
number of authors 
per paper") + 
  theme_classic(base_size = 14)+theme(panel.grid.major=element_line(colour="grey90"))

p


b <- dplyr::group_by(pubdata_art, Publication.Date) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(Category.Normalized.Citation.Impact, na.rm = TRUE),
    median = median(Category.Normalized.Citation.Impact, na.rm = TRUE),
    sd = sd(Category.Normalized.Citation.Impact, na.rm = TRUE)
  )


p <- ggplot(b, aes(Publication.Date, y=mean)) + geom_point(color = "black") + 
  ylim(0,6) + xlim(1995,2020) + labs(x="Year", y = "mean, 
Category Normalized 
Citation Impact") +
  theme_classic(base_size = 14)+theme(panel.grid.major=element_line(colour="grey90"))
p
```

#Session information
```{r}
sessionInfo()
```

